---
title: "1_BasicR"
author: "RNepal"
date: "9/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Syntax

* Assign value to a variable using <- operator
* Assigned a number to the variable but  can assign many type of data types
```{r Basic syntax}
num <- 123
print(num)

alp <- "abc"
print(alp)
```

# Comment

* Comments are  helping text in  R program
* Ignored by the interpreter while executing your actual program. 
* Written using .green[#] in the begining of the statement. 
* Write comments as it help to make code readable ,reuseable.


```{r}
# assign 1 to a
a <- 1
a
```


# Your Turn 1

* Assign some number to num1 variable and print out.
* Assign alphabet to alpha1 variable and print out.
* Dont forget to write some comments.

```{r}
# Assign number to num1
num1 <- 3
class(num1)

# Assign number to alpha1

string <- "suman"
class(string)
```


# Variable Naming convention

```{r}
my_data <- 234 # correct
my_data_1 <- 456
1_my_data <- 345 # incorrect
```

#Variable names are case sensitive
#Example below mydata and MyData are different

```{r}
mydata <- 234
MyData <- 567 
print(mydata)
print(MyData)
```

# Data Types

* types or kind of information or data a variable is
holding
* class() function is used to determine the datatype of variable 

Logical/Boolean : TRUE /FALSE 
Numeric : 123, 23.4, 345
Character: "a","abc","This is a ball."

# Data Types

```{r}
class(a)
```

```{r}
class(alp)
```

```{r}
class(mydata)
```

# Your Turn 2

* Find the data type of the variable you defined in Your Turn 1

# Data Structures
Data Structure are the R Objects that is used to store some kind of data.

* Vectors
* Lists
* Matrices
* Arrays
* Factors
* Data Frames

#Insert chunk 
#Chunk is inserted with shortcut Ctrl + Alt + I





# Vector

* Vector is collection of same type of values
* created using c() function

```{r}
# create new vector

vec1 <- c(1,2,3)
vec1
```

```{r}
class(vec1)
```

```{r}
vec2 <- c("ram","shyam",1)
vec2
```

```{r}
class(vec2)
```


# Your Turn 3

* create a new vector
* find the datatype of vector

```{r}
# create a new vector 
new_vec <- c(1,2,3)

# find the class of vector
class(new_vec)
```


# Operators

## Arthmetic Operators

Operator | Meaning
-------- | -------
+ or - | Add or Subtract
* | Multiplication
/ | Division
^ | Exponent
%% | Modulus

## Addition
```{r Addition}
#Addition
a <- c(5,6,7)
b <- c(2,1,3)
print(a - b)
```


## subtraction
```{r subtraction}
#subtraction
a <- c(9,8,7)
b <- c(2,4,9)
print(a - b)
```

## Multiply
```{r Multiplication}
# Multiply
a <- c(1,2,3)
b <- c(2,3,4)
print(a*b)
```

## Division
```{r Division}
# Division
a <- c(22,13,24)
b <- c(2,3,4)
print(a/b)
class(a/b)
```

## Exponent
```{r}
4 ^ 2
```

```{r}
99 ^ 3
```

## Modulus

```{r}
7 %% 2
```
```{r}
99 %% 8
```

# Relational Operators

* operators  used to compare two elements or vectors.
* compare vector a with vector b.



Operator | Definition
-------- | ---------
==       | Equal to  
>=       | Greater than or equal to
<=       | Lesser than or equal to
>        | Greater than
<        | Lesser than
!=       | Not equal to

# Greater than ( > )
```{r}
# greater than

a = c(5,2,3, 9)
b = c(10,1,2, 4)
print(a > b)
```



# Less Than ( < )
```{r}
#less than 
a <- c(9,10,11,21)
b <- c(11,2,3,55)
print(a<b)
```

# Equals to ( == )
```{r}
#equals to

a <- c(2,3,4)
b <- c(2,3,5)
print(a == b)
```

# NotEqual ( != )
```{r}
# notequal to
a <- c(2,3,4,10)
b <- c(1,3,4,9)
print(a!=b)
```


# Less than equal to
```{r}
#less than equal to
a <- c(2,3,4,9)
b <- c(2,5,7,8)
print(a<=b)
```


# Greater than equal to
```{r}
# greater than equal to

a <- c(2,3,4,10)
b <- c(5,7,4,9)
print(a>=b)
```


# Logical Operators

# AND operator
```{r}
# AND operator

a <- c(3,9,5)
b <- c(5,6,7)
print (a>5)
print(b>5)
print(a>5 & b>5)
```


# OR operator

```{r}
# OR operator

a <- c(1,5,3)
b <- c(4,5,6)
print(a<5)
print(b>5)
print(a<5 | b>5)

```


# Logical NOT ( ! )

```{r}
# logical not

a <- c(3,4,5)
print(a>3)
print(!(a>3))
```


# Data Frame

* table with rows and columns.
* each column contains values of one variable
* each row contains set of values from each column

```{r}
head(mtcars)
```


# lets Create dataframe

You can create a dataframe using different type of vector in r using data.frame()

R by default converts string to factor while reading or creating dataframe, to keep string as string we use stringsAsFactors.


```{r}
data <- data.frame(id = c(1:5),
                   name = c("Richa","Deepak","Manisha","Ryan","Ram"),
                   age = c(23,25,45,30,57),
                   gender=as.factor(c("female","male","female","male","male")),
                   stringsAsFactors = FALSE)
```

```{r}
data
```

# Get the Structure of the Data Frame
```{r}
# Get the structure of the data frame.
str(data)
```

# Summary of Data in Dataframe

```{r}
# Print the summary
library(psych)
summary(data)
describe(data)
```

# Extract Data From DataFrame

## head
* head() extract first 6 rows of the dataframe

```{r}
head(mtcars)
```

```{r}
head(mtcars,4)
```


## tail

* tail() extract last 6 rows of the dataframe

```{r}
tail(mtcars)
```


```{r}
tail(mtcars,10)
```

# Extract Data

* dataframe_name[row_id_from: row_id_to, col_id_from:col_id_to]

## Extract first two rows and first two columns

```{r echo=FALSE}
# extract first two rows and column
data[1:2,1:2]
```

## Extract 2nd row and all columns

```{r}
data[2,]
```

## Extract 3rd   row with 2nd and 4th column

```{r}
data[3,c(2,4)]
```



## Extract 2nd  and 4th row with 2nd, 3rd and 4th  column

```{r}

data[c(2,4),c(2:4)]
```

# your turn 4

* Extract name and age column
* Extract 2nd row  
* Extract 3rd row and name, gender,age column

## Extract name and age column
```{r}
data
data[c("name","age")]
```

## Extract 2nd row  
```{r}
data[2,]
```

## Extract 3rd row and name, gender,age column

```{r}
data[3, c("name", "gender", "age")]
```



# Add Column

```{r}
#Add the 'address' column
data$address <- c("butwal","kathmandu","birgunj","ilam","karnali")
data
```



# Your Turn 5

*  Add blood_group column to dataframe

```{r your turn 3}
data$blood_group <- as.factor(c("A", "B","AB","O", "A"))
data
```

# Add Row

```{r}
# Create second data frame
newdata <- data.frame(id = c(6:8),
                   name = as.character(c("Deepika","Manish","Brion")),
                   age = c(33,15,40),
                   gender=as.factor(c("female","male","male")),
                   blood_group=c("A", "B", "AB"),
                   address=c("patan","dharan","butwal"),
                   stringsAsFactors = FALSE)
# connect two dataframe with rows
final_data <- rbind(data,newdata)
```

```{r}
final_data
```

# Merge Two dataframe

* Connect two dataframe with different columns using a .green[key]
* .merge() is used to merge the two dataframe
* must have a common column in two dataframe


```{r}
prof_data <- data.frame(
                   name  = as.character(c("Deepika","Manish","Brion","Richa","Deepak","Manisha","Ryan","Ram")),
                   phone  = c(9129345550,9129365570,9129346789,9129343457,9129342344,9129343451,9129344670,9129345650),
                   
profession =  as.character(c("Doctor","Nurse","Driver","Scientist","Computer Technician","Teacher","Student","Lawyer")),
                   stringsAsFactors = FALSE)
prof_data
```


```{r}
# merge prof_data and final_dat using name by key
merge(prof_data,final_data, by = 'name')

```

```{r}
merge(final_data, prof_data, by = 'name')
```



# Rename Column

* select column with its name and replace with new name
* name() function gives the name of the dataframe column
```{r}
names(final_data)[names(final_data)=="name"] <- "first_name"
final_data
```


# Function

* set of statements organized together to perform a specific task
* R has a large number of in-built functions

```{r}
# function for adding number
add_me <- function(num1,num2){
  sum <- num1 + num2
  return(sum)
}
```

```{r}
add_me(99,123)
```


# Built In Function

```{r}
# Create a sequence of numbers from 32 to 44.
print(seq(32,44))

# Find mean of numbers from 25 to 82.
print(mean(25:82))

# Find sum of numbers frm 41 to 68.
print(sum(41:68))
```

# Packages

 * Packages are a collection of R functions, compiled code and sample data
 * Stored under a directory called "library" in the R environment
 * R installs a set of packages during installation 
 * Comprehensive R Archive Network (CRAN) has collection of packages
 
 
#  Asking for Help

* ? and .help() function  is used to ask help
* help provides details  about  r objects like function, data, package.

```{r}
# asking help about the mean() function
?mean
```

```{r}
# asking help about the mtcars datasets
?mtcars
```

```{r}
# asking help about the dplyr package
library(dplyr)
?dplyr
```

