---
title: "Data Transformation"
author: "RNepal"
date: "9/16/2019"
output: html_document
---


<!-- This file is provided in Data Analysis with R bootcamp by R Nepal (www.rugnepal.org) -->

## Load library

```{r setup, include = FALSE}
library(haven)
library(readxl)
library(readr)
library(writexl)
```

## Load Data

```{r}
library(readr)

gapminder <- read_csv("./data/gapminder.csv")

gapminder

```

## Select - Variables

```{r}
## %>%  <------- Ctrl + Alt + m  (Keyboard shortcut)

country_gdp_year <- gapminder %>% select(year, country, gdpPercap)
  
  
# Displays data, so don't remove them below as well
country_gdp_year
```

# Filter by value

```{r}
gapminder_asia <- gapminder %>%
  filter(continent == "Asia") 
  

gapminder_asia
```


## Your Turn - 01

Display country-wise, annual life expectancy of Europe.

```{r}
lifeexp_year_europe <-  gapminder %>%
  filter(continent=="Europe") %>% select(year,country, lifeExp)
  


# displays data
lifeexp_year_europe
```


# Mutate - Add New Variables

```{r}
gapminder_total_gdp <- gapminder %>%
  mutate(gdp = gdpPercap*pop )

# displays data 
gapminder_total_gdp
```

# Your Turn - 02

Find each countries in Asia with life expectancy in 2007 and rank of country's life expectancy.

```{r}
library(dplyr)
asia_lifeExp <-  gapminder %>% filter(continent=="Asia", year==2007) %>%
  select(country, lifeExp) %>% mutate(rank=min_rank(lifeExp))
   
  

asia_lifeExp  
```

# Group By - One or more variable

```{r}
gapminder_group_by <- gapminder %>%
  filter(year == 2007 ) %>%
  group_by(continent)
  
gapminder_group_by
```

# Arrange - Ascending Order

```{r}
gapminder_asc <- gapminder %>%
  filter(year == 2007 ) %>%
  select(country, pop) %>% 
  arrange(pop)
  
  

gapminder_asc 
```

# Arrange - Descending Order

```{r}
gapminder_desc <- gapminder %>%
  filter(year == 2007 ) %>%
  select(country, pop) %>% 
  arrange(desc(pop))
  
  
# displays data
gapminder_desc 
```


# Your Turn - 03

Find countries in Asia with low ranked gdpPercap in 1992

```{r}
asia_gdpPercap_1992 <- gapminder %>% filter(continent=="Asia", year==1992) %>% 
  select(country, gdpPercap) %>% 
  mutate(Rank=min_rank(gdpPercap)) %>% arrange(Rank)
  
  
asia_gdpPercap_1992 
```

# Summarize Data

```{r}
gampinder_sum <- gapminder %>% filter(year==2007) %>% group_by(continent) %>% summarise(MeanLifeExp=mean(lifeExp), gdp=mean(gdpPercap))

  
gampinder_sum
```


