---
title: "Visualization with ggplot2"
author: "RNepal"
date: "9/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Grammar of Graphics



 ggplot2 is a data visualization tool based on the grammar of graphics and using different functions.


## "The simple graph has brought more information to the data analyst's mind than any other device." - John Tuckey

# Setup

```{r setup}
library(ggplot2)
library(readr)
library(dplyr)
```

# Data
```{r}
bechdel <- read_csv("./data/bechdel.csv")
```

```{r}
head(bechdel)
```


The raw data behind the story "The Dollar-And-Cents Case Against Hollywood's Exclusion of Women" from the fivethirtyeight .
There are 15 columns in the dataset and  are described in the section below.


* year :Year of release

* imdb :Text to construct IMDB url. Ex: https://www.imdb.com/title/tt1711425

* title :Movie title

* test :bechdel test result (detailed, with discrepancies indicated)

* clean_test : bechdel test result (detailed): ok = passes test, dubious, men = women only talk about men, notalk = women don't talk to each other, nowomen = fewer than two women

* binary :Bechdel Test PASS vs FAIL binary

* budget :Film budget



* domgross :Domestic (US) gross

* intgross :Total International (i.e., worldwide) gross

* code :Bechdel Code

* budget_2013 :Budget in 2013 inflation adjusted dollars

* domgross_2013 :Domestic gross (US) in 2013 inflation adjusted dollars

* intgross_2013 :Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

* period_code
* decade_code



# Your Turn 1

```{r}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget,
                          y = domgross))
```

# ggplot 2 Template

##Format of ggplot2 
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))


# Aesthetics

* something we can see in graph
* X and Y axis, color, size ,shape of the graph

## color 
```{r}
# color is set to clean_test
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, 
                          y = domgross, 
                          color = clean_test))
```

## shape
```{r}
# shape is set to clean_test
ggplot(bechdel) +
  geom_point(aes(x = budget,
                 y = domgross,
                 shape = clean_test))
```

# Your Turn 2

* Experiment adding color, size , alpha , and shape aesthetics to your graph

* How do aesthetics behave different when mapped to discrete and continous variable?

* What happens when you use more than one aesthetic?

```{r}
# color aesthetic
ggplot(data = bechdel) +
    geom_point(mapping = aes(x = budget , y = domgross, color = clean_test))

```


```{r}
# shape aesthetic
ggplot( data = bechdel) +
    geom_point(mapping = aes(x = budget, y = domgross, shape = clean_test))
```


```{r}
# alpha aesthetic 
ggplot( data = bechdel) +
    geom_point(mapping = aes(x = budget, y = domgross, color = clean_test), alpha = 0.2)
```





```{r}
ggplot(data = bechdel) +
  geom_point(mapping = aes(x = budget,
                           y = intgross),
             color ="red", alpha=0.2)
```

# Geoms

* functions that define the geometric or visual  object  of the graphs.

* used to define different type of plots like bar charts, boxplots, line charts and many more.


Geom | Description
--- | ---
geom_point() | creates scatterplot
geom_boxplot() | creates boxplot
geom_histogram() | creates histogram
geom_density() | creates density plot
geom_bar() | creates barplot
geom_line()| creates lineplot

]

## Scatterplot

* Scatterplot is used to define the relationship between two variable 
* geom_point() geom is used for scatterplot.

```{r}
ggplot(data = bechdel) +
  geom_point(aes(
    x = budget,
    y = domgross),
    alpha = 0.3,color="red")
```

## Bubble Chart


* bubblechart is same like scatterplot with points being bubbles
* bubblechart is defined using geom_point() geoms


```{r message=FALSE, warning=FALSE}
ggplot(data = bechdel) +
  geom_point(aes(
    x = budget,
    y = domgross,
    size = .001*domgross
  ),
    alpha = 0.3,color="red")
```

## Boxplot

* boxplot can be defined using geom_boxplot() geoms

* used to define .green[quartiles] and find outliers.

```{r}
ggplot(data = bechdel) +
  geom_boxplot(aes(x = clean_test,
                   y =  budget))
```


#Violin plot in ggplot2

```{r}
ggplot(data = bechdel) +
  geom_violin(aes(x = clean_test,
                   y =  budget))
```


# Your turn 3

### scatterplot of the domgross and intgross

```{r}
ggplot(data = bechdel) +
  geom_point(aes(x = domgross,
                             y = intgross))
```

### boxplot of the domgross VS  binary

```{r}
ggplot(data = bechdel) +
  geom_boxplot(aes(x = binary,
                             y = domgross))
```

## Histogram

* shows the distribution of data
* geom_histogram() geom is used.


```{r}
ggplot(data = bechdel) +
  geom_histogram(aes(x = budget ))
```


* binwidth gives width to bins.

```{r}
ggplot(data = bechdel) +
  geom_histogram(aes(x = budget ), 
                 binwidth = 10000000)
```



```{r echo=FALSE}
ggplot(data = bechdel) +
  geom_boxplot(mapping = aes(y = intgross ))
```


# Your Turn 4

* Make the Histogarm  of intgross 

```{r echo=FALSE}
ggplot(data = bechdel) +
  geom_histogram(mapping = aes(x = intgross ))
```

## Density Plot

* shows the distribution of data
* geom_density() geom is used for density plot

```{r}
ggplot(data = bechdel) +
  geom_density(mapping = 
                   aes(x = budget))
```

# Your Turn 5

* Make the Density  of budget colored by clean_test <br> as shown below
 
```{r}
ggplot(data = bechdel) +
  geom_density(mapping = aes(x = budget  ,color = clean_test  ))
```
 
# Bar Plot

* visualize the count of the values
* geom_bar() geom is used for barplot.

```{r}
ggplot(data = bechdel) +
  geom_bar(aes(x = clean_test))
```

# Your Turn  6

* Make the bar chart of clean_test and colored by clean_test as shown below.

```{r echo=FALSE}
ggplot(data = bechdel) +
  geom_bar(mapping = aes(x = clean_test,
                         fill = clean_test))
```

# Coordinate scales

* Arrange the width of the bar


```{r}
ggplot(bechdel, aes(x = clean_test , 
                  y = domgross)) +
  geom_col(aes(fill = clean_test),
           width = 0.4)


```

# Your turn 7

*  Add a width adjustment to this plot
```{r}
ggplot(bechdel, aes(x = binary ,
                  y = budget)) +
  geom_col(aes(fill = binary), width =0.4 )
```

# Pie Chart

* pie chart is created rotating the barchart
* pie chart round shaped 

```{r}
ggplot(data = bechdel,aes(x = factor(1), 
                        fill = binary))+
  geom_bar(width = 1)
```

```{r}
ggplot(data = bechdel,aes(x = factor(1),
                        fill = binary))+
  geom_bar(width = 1) + coord_polar("y")
```

```{r}
ggplot(data = bechdel,aes(x = factor(1), 
                        fill = binary))+
  geom_bar(width = 1) +
  coord_polar("y")+theme_void()
```


```{r}
# create new dataframe
binary_data <- bechdel %>%
  count(binary) %>%
  mutate(percent = (n / sum(n)) * 100)
ggplot(binary_data, aes(x = "", y = percent, fill = binary)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
# convert barplot to polar coordinates     
  coord_polar("y", start = 0) +
# add labels 
  geom_text(aes(label = paste0(round(percent, 2), "%")), position = position_stack(vjust = 0.5), color = "white") +
# add color scale manually
  scale_fill_manual(values = c("#55DDE0", "#F6AE2D")) +
  theme_void()
```

# Lineplot

```{r}
bechdel %>% count(year)%>%ggplot() +
 geom_point(mapping = aes(x = year, y = n)) +
 geom_line(mapping = aes(x = year, y = n))
```


```{r}
bechdel %>% count(year)%>%ggplot() +
 geom_point(mapping = aes(x = year, y = n),color = "red") +
 geom_line(mapping = aes(x = year, y = n),color="green")
```

# Title , Axis and caption

* labs() is used to add title,caption etc
```{r}
ggplot(data = bechdel,aes(x = binary))+
    geom_bar(aes(fill = binary)) + 
    labs(title = "Graph title", 
         x = "X-axis", 
         y = "Y-axis",
         caption = "data source: xyz")
```


# Your Turn 8

* Add  Title,caption to belwo plot 

```{r}
ggplot(data = bechdel,aes(x = clean_test))+
  geom_bar(aes(fill = clean_test)) +
  labs(title = "Bar plots", 
         x = "Clean_test", 
         y = "Count",
         caption = "Data source: fivethirtyeight")
```

# Change axis tick mark labels

```{r}
ggplot(data = bechdel,aes(x = test))+
  geom_bar(aes(fill = test))
```


```{r}
ggplot(data = bechdel,aes(x = test))+
  geom_bar(aes(fill = test)) +
    theme(axis.text.x = element_text(angle = 90))
```

# Flip the plot
```{r echo=TRUE}
ggplot(data = bechdel,aes(x = test))+
  geom_bar(aes(fill = test)) +coord_flip()
```

# Your Turn 9
* Rotate the x axis ticks by 45 degree angle.
```{r}
ggplot(data = bechdel,aes(x = test))+
  geom_bar(aes(fill = test)) +
    theme(axis.text.x = element_text(angle = 45))
```


# Legend

## Remove legend

```{r}
ggplot(data = bechdel) +
  geom_bar(mapping = aes(
    x = clean_test,
    fill = clean_test
  )) +
  guides(fill = FALSE)
```


# legend position

```{r}
ggplot(data = bechdel) +
  geom_bar(aes(x = clean_test,
               fill = clean_test)) +
  theme(legend.position="top")
```

## Change legend title

```{r}
ggplot(data = bechdel) +
  geom_bar(mapping = aes(x = clean_test,
                         fill = clean_test))+
    labs(fill = "Clean Test", title="Hello plot")
```


# Labels

* geom_text() adds text directly to the plot.
* geom_label() create a a rectangle behind the text. 

```{r}
bechdel %>% count(clean_test) %>% ggplot() +
    geom_col(aes(x=clean_test,y=n,fill= clean_test)) +
    geom_label(aes(x= clean_test, y = n, label = n)) +
    labs(title = "Barplot of clean test", y = "Count")
```


# Arrange the barplot

* reorder() arrange the values 

```{r}
bechdel %>% count(clean_test) %>% ggplot() +
    geom_col(aes(x=reorder(clean_test,n),y=n,fill= clean_test)) +
    geom_label(aes(x= clean_test, y = n, label = n)) +
    labs(title = "Distribution of clean test", y = "Count",x = "")
```


```{r}
bechdel %>% count(clean_test) %>% ggplot() +
    geom_col(aes(x=reorder(clean_test,-n),y=n,fill= clean_test)) +
    geom_label(aes(x= clean_test, y = n, label = n)) +
    labs(title = "Distribution of clean test", y = "Count",x = "")
```



```{r}
bechdel %>% count(clean_test) %>% ggplot() +
    geom_col(aes(x=reorder(clean_test,n),y=n,fill= clean_test)) +
    geom_label(aes(x= clean_test, y = n, label = n)) +
    labs(title = "Distribution of clean test", y = "Count",x = "")+coord_flip()
```




```{r}
bechdel %>% count(clean_test) %>% ggplot() +
    geom_col(aes(x=reorder(clean_test,-n),y=n,fill= clean_test)) +
    geom_label(aes(x= clean_test, y = n, label = n)) +
    labs(title = "Distribution of clean test", y = "Count",x = "")+coord_flip()
```




# Position

* change the position of the graph plots

```{r}
bechdel %>% filter(year == c(2013,2012,2011,2010,2009))%>%
ggplot(aes(x = year)) +
 geom_bar(aes(fill = binary))
```

```{r}
bechdel %>% filter(year == c(2013,2012,2011,2010,2009))%>%
ggplot(aes(x = year)) +
 geom_bar(aes(fill = binary), position = 'dodge') 
```

# Scales

* assign different color palette
* Enter scale_color_brewer

```{r}
?scale_color_brewer
```



```{r echo=FALSE}
ggplot(data = bechdel, aes(x = clean_test)) +
  geom_bar(aes(fill = clean_test)) +
  scale_fill_brewer(palette = "Set2")
```




```{r}
ggplot(data = bechdel, aes(x = clean_test)) +
  geom_bar(aes(fill = clean_test)) +
  scale_fill_brewer(palette = "Set1")
```

# Your turn 8

*  Pratice with different color palette 

```{r}
ggplot(data = bechdel, aes(x = clean_test)) +
  geom_bar(aes(fill = clean_test)) +
  scale_fill_brewer(palette = "Pastel1" )
```

# Add color manually

```{r}
ggplot(data = bechdel, aes(x = clean_test)) +
  geom_bar(aes(fill = clean_test)) +
  scale_fill_manual(values = c("#55DDE0", "#33658A", "#2F4858", "#F6AE2D", "#F26419"))
```

# Facets

* Layout panel for the graphs
* facet_wrap() & facet_grid()


```{r message=FALSE, warning=FALSE}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, y = domgross, color = clean_test))
```

```{r}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, y = domgross, color = clean_test)) +
 facet_wrap(~clean_test)
```


```{r}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, y = domgross, color = clean_test)) +
 facet_grid(period_code ~ clean_test)+theme(axis.text.x = element_text(angle=90))
```

# Themes

### theme_bw()

```{r echo=TRUE}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, y = intgross, color = clean_test)) +
    theme_bw()
```


### theme_classic()

```{r echo=TRUE}
ggplot(data = bechdel) +
 geom_point(mapping = aes(x = budget, y = intgross, color = clean_test))+theme_classic()
```


# Your turn 9

* Try different theme for this graph

```{r echo=TRUE, eval=FALSE,include=TRUE}
ggplot(data = bechdel,aes(x = clean_test)) +
  geom_bar(aes(fill = clean_test)) + labs(title="Hello world chart", x="clean_test", y="count")+theme_classic()
```


#  Saving plots

* ggsave() is used to save plot.

```{r}
# assign the ggplot graph to myplot variable

myplot <- bechdel %>% count(clean_test) %>% ggplot() +
# plot barplot
    geom_col(aes(x=reorder(clean_test,n),y=n,fill= clean_test)) +
# adding labels to plot
    geom_label(aes(x= clean_test, y = n, label = n)) +
# adding title and edit xaxis and y axis
    labs(title = "Distribution of clean test", y = "Count", x= "",fill="Clean Test")+
    scale_fill_brewer(palette = "Set1") +
# use theme minimal
    theme_classic()
myplot
```


```{r}
myplot <- bechdel %>% count(clean_test) %>% ggplot() +
# plot barplot
    geom_col(aes(x=reorder(clean_test,n),y=n,fill= clean_test)) +
# adding labels to plot
    geom_label(aes(x= clean_test, y = n, label = paste0(round(n/sum(n)*100,2),"%") )) +
# adding title and edit xaxis and y axis
    labs(title = "Distribution of clean test", y = "Count", x= "",fill="Clean Test")+
    scale_fill_brewer(palette = "Set2") +
# use theme minimal
    theme_classic() + scale_y_continuous(expand = c(0, 0),limits = c(0,850))
myplot
```



```{r}
# saves the plot to working directory
ggsave("myplot.png", plot = myplot , width =  6, height = 6, units = "in", dpi = "retina")
```

