
title: "03-Statistics"
author: "RNepal"
date: "9/22/2019"
output: html_document


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#setup

```{r}
library(readr)
library(tidyverse)
heart <- read_csv("./data/heart_disease.csv")
```

# About Data

* slope_of_peak_exercise_st_segment (type: int): the slope of the peak exercise ST segment, an electrocardiography read out indicating quality of blood flow to the heart

* thal (type: categorical): results of thallium stress test measuring blood flow to the heart, with possible values normal, fixed_defect, reversible_defect

* resting_blood_pressure (type: int): resting blood pressure

* chest_pain_type (type: int): chest pain type (4 values)

* num_major_vessels (type: int): number of major vessels (0-3) colored by flourosopy

* fasting_blood_sugarl (type: binary): fasting blood sugar > 120 mg/dl

* resting_ekg_results (type: int): resting electrocardiographic results (values 0,1,2)

* serum_cholesterol (type: int): serum cholestoral in mg/dl

* oldpeak_eq_st (type: float): oldpeak = ST depression induced by exercise relative to rest, a measure of abnormality in electrocardiograms

* sex (type: binary):  female, male

* age (type: int): age in years

* max_heart_rate_achieved (type: int): maximum heart rate achieved (beats per minute)

* exercise_induced_angina (type: binary): exercise-induced chest pain (No, Yes)

# Statistical Terms

* Population: total set of observations

* Sample: Portion of Population

* Observations: measure or count during a study or experiment

* Variables: characteristics or quantity that can be counted 







# Types of Data

* Numeric : have discrete or continous data (1,2,3,4,5)
* Categorical: data are categories or non-numeric data (TRUE/FALSE, low/middle/high )


# Type of Stat

* Descriptive 
* Inferential


# Descriptive Stat

* summarizes the data
* focuse on distribution, central tendancy and dispersion of the data


# Central Tendancy 
A measure of central tendency is a single value that attempts to describe a set of data by identifying the central position within that set of data.

* Mean
* Median
* Mode


# Mean

* Average of the given data

```{r average of the patient}
mean(heart$age)
```



# Median

* Middle point of the data
* 50 percentile of the data 
```{r median age of the patient}
median(heart$age)
```


# Mode

* Value with highest frequency
```{r mode of the thal}
# function for finding mode value
mode_find <- function(data){
  # counts the value
  count <- table(data)
  # find the value with maximum frequency
  mode_value <- names(count)[which(count == max(count))]
  return(mode_value)
}
mode_find(heart$thal_test)
```


# Your Turn 1

* find the mean and median of the serum_cholesterol
* find the mode of chest_pain_type



```{r}
mean(heart$serum_cholesterol)
```

```{r}
median(heart$serum_cholesterol)
```


```{r}
mode_find(heart$chest_pain_type)
```


class:top, left
# Range


* Difference between the largest and smallest points in the data
* range()  is used.

```{r echo=TRUE}
range(heart$age)
```


* min() and max() can also be used to find Range.

```{r range of the max_heart_rate, echo=TRUE}
min(heart$max_heart_rate)
max(heart$max_heart_rate)
```





# Interquartile range

* Measure of difference between 75 Percentile (3rd quartile) and 25 Percentile ( 1st quartile)
* IQR() is used 

```{r IQR of the Patient age}
IQR(heart$age)
```


# Quartile

* Quartiles are the values that divide a list of numbers into quarters.
* quantile() is used

```{r}
quantile(heart$resting_blood_pressure)
```


# Your Turn 2

* find range and IQR of the serum_cholesterol
* find range and IQR of the max_heart_rate


```{r}

range(heart$serum_cholesterol)
```

```{r}

IQR(heart$serum_cholesterol)
```



 range and IQR of the max_heart_rate
```{r}

range(heart$max_heart_rate)
```

```{r}
print("inter quartile range:")
IQR(heart$max_heart_rate)
```



# Variance 

* Average of squared difference from the mean 
* Used to measure the spreadness of the data
* var() is used

```{r variance of the resting_blood_pressure}
var(heart$resting_blood_pressure)
```


# Standard Deviation

* Square root of variance & measures spread of the data
* Easier to apply than variance 
* sd() is used.

```{r standard deviation of the resting_blood_pressure}
sd(heart$resting_blood_pressure)
```


# Your Turn 3

* find variance and standard deviation of age
* find variance and standard deviation of serum_cholesterol




# variance and standard deviation of age

```{r}
var(heart$age)
sd(heart$age)
```


## variance and standard deviation of serum_cholesterol
```{r}
var(heart$serum_cholesterol)
sd(heart$serum_cholesterol)
```



# Normal Distribution

* symmetric distribution where most of the observations cluster around the central peak
* Bell Shaped curve/ Gaussian distribution





# Normal Distribution

```{r Histogram of the patient age,message=FALSE, warning=FALSE}
ggplot(heart,aes(x = age))+
  geom_histogram()
```


```{r Density Plot of the patient age,message=FALSE, warning=FALSE}
ggplot(heart,aes(x = age))+
  geom_density()
```

# Normal Distribution

```{r}
qqnorm(heart$age)
qqline(heart$age)
```

# Shapiro Test

* p-value more than 0.05 conclude data is normally distributed.

```{r}
shapiro.test(heart$age)
```

* age is normally distributed


# Your Turn 4

* Is resting_blood_pressure Normally Distributed?
* Is max_heart_rate Normally Distributed?


```{r}
ggplot(heart,aes(x = resting_blood_pressure ))+
  geom_density()
```



```{r}
qqnorm(heart$resting_blood_pressure)
qqline(heart$resting_blood_pressure)
```

```{r}
shapiro.test(heart$resting_blood_pressure)
```

# Skewness

* measure of how symmetric a distribution is
* how much the distribution is different from the normal distribution
* moderately skewed: -1 to -0.5 and 0.5 to 1
* Approximately symmetric: -0.5 to 0.5

```{r}
#library(moments)
#skewness(heart$max_heart_rate)
```

The max_heart_rate is skewed to the left.




# Kurtosis
* measure that defines how heavily the tails of a distribution differ from the tails of a normal distribution
* define whether the tails of a given distribution contain extreme values

```{r}
#kurtosis(heart$age)
```

The distribution of age is platykurtic as the  computed value is less than 3.

# summary()  and str()

* summary() gives the basic statistic of the data.
* str() gives the structure of the data.





### Summary
```{r}
summary(heart)
```



### str

```{r}
str(heart)
```





# Inferential Statistics

* Describes and make inferences about the population and the sampled data
* hypothesis testing
* estimating of parameters


# Hypothesis testing

* a research
question is a hypothesis asked in question format
* question : "Does Women multitask better than men? "
* hypothesis : "There
is a significant difference between the women and men in multitasking."
* evaluates two mutually exclusive statements


# Null and Alternate Hypothesis
* Null Hypothesis : "Women does not multitask better than men."
* Null Hypothesis denoted by Ho


* Alternate Hypothesis :"Women multitask better than men."
* Alternate Hypothesis denoted by H1

* if p-value is greater than 0.05 ,accept null hypothesis.
* smaller p-value means that there is stronger evidence in favor of the alternative hypothesis


### Types of Errors in Hypothesis Testing

* type I error is a rejection of the null hypothesis when it is really true.
* type II error is a failure to reject a null hypothesis that is false



# T- Test 

* used to determine whether the mean between two data points or samples are
equal to each other

### One Sample T-Test

* one-sample t-test is used to test whether the mean of a population is
equal to a specified mean
* Question :Is the mean of the Resting blood pressure is equal to specified mean suppose 120?
* Null Hypothesis : There is no difference between the resting blood pressure  mean  and specified mean.
* Alternative Hypothesis :  There is significant difference between the mean of the resting blood pressure mean nd specified mean.


```{r}
ttest_max_heart <-t.test(heart$resting_blood_pressure, mu = 120)
ttest_max_heart
```

As the p-value is less than 0.05 threshold . We can reject null hypothesis at 95% confidence interval. 
There is significant difference between the mean resting blood pressure and the specified one.


# Your Turn 5
* Is the mean of the Max Heart Beat is equal to specified mean suppose 90?

```{r}
t.test(heart$max_heart_rate , mu = 90 )
```


# Two-Sample Independent T-Test


*  two compare two means of mutually exclusive group( Male / Female)

* Is there significant difference between the mean of  male  age and mean of female age ?
* Null: There no significant difference between the mean of male age and female age.

* Alternate: There is significant difference between the mean of male age and female age.


```{r}
# extract the male and female data 
male_max<- heart %>% filter(sex == "male")
female_max <- heart %>% filter(sex == "female")
t.test(male_max$age,female_max$age, var.equal = TRUE, paired = FALSE)
```


* As the  p-value = 0.4719 is greater than 0.05 .We accept the null hypothesis.
* There no significant difference between the mean of .blueb[male age] and .pink female age.


# Your Turn 6

* Compare mean male and female max_heart_rate
```{r}
male_max_heart<- heart %>% filter(sex == "male")
female_max_heart <- heart %>% filter(sex == "female")
t.test(male_max_heart$max_heart_rate,female_max_heart$max_heart_rate, var.equal = TRUE, paired = FALSE)
```


# Two Sample Dependent T-Test

* Used  to  compare two means of two dependent samples.
* Any effect or difference before interventional program and after interventional program.

```{r}
# creating random data 
set.seed(123);
before <- rnorm(100, mean=2, sd=1);
after <- rnorm(100, mean=3, sd=1);

data <- data.frame(before, after);
t.test(data$before, data$after, paired=TRUE);
```

* As the  p-value = 0.4719 is greater than 0.05. We reject the null hypothesis.
* There is  significant difference between the mean of before and after intervation.

# Chi Square test
* used to compare the relationships between two categorical variables

# Contingency Test

* Used to compare relationship between two categorical variables
* Is there association between chest_pain_type and heart_disease ?
* null: There is no association between chest_pain_type and heart_disease.
* alternate: There is association between chest_pain_type and heart_disease.

```{r}
chisq_chest_heart <- chisq.test(heart$chest_pain_type, heart$heart_disease)
chisq_chest_heart
```

```{r}
library(broom)
library(knitr)
chisq_chest_heart$observed %>% kable()
```

If the the value in any cell of the observation is less than 5 equal to 5. Use Fisher test.

```{r}
fisher.test(heart$chest_pain_type, heart$heart_disease)
```

As, p-value is less than 0.05 , Null hypothesis is rejected.There is association between chest_pain_type and heart_disease.




# Your Turn 7
* Is there association between sex and heart_disease?



```{r}
chisq.test(heart$sex, heart$heart_disease)
```




# Correlation

* define how close two variables are 
* range between (-1 to +1)


```{r correlation of the patient age and heart_disease}
cor(heart$age,heart$max_heart_rate)
```



# Your Turn 8

* find the correlation between resting_blood_pressure and max_heart_rate
* find the correlation between serum_cholesterol and age

```{r}
cor(heart$resting_blood_pressure, heart$max_heart_rate)
```


```{r}
# correlation between resting_blood_pressure and max_heart_rate
cor(heart$resting_blood_pressure,heart$max_heart_rate)
```

```{r}
# correlation between serum_cholesterol and age
cor(heart$serum_cholesterol, heart$age)
```

```{r}
ggplot(heart) + geom_point((aes(x=age, y=max_heart_rate))) + geom_smooth(aes(x=age, y=max_heart_rate),method=lm,se=FALSE)
```

